Argennon: A Scalable Smart Contract Platform

### 2021-06-28 tex/sections/consensus.tex
Add changes needed by the recovery protocol
Signed-off-by: aybehrouz <behrouz_ayati@yahoo.com>

There are two primary types of certificate committees in Argennon: the committee of \emph{delegates} and the committee
of \emph{validators}. Argennon has one committee of delegates and $m$ committees of validators.

The committee of delegates generates a certificate for every block of the Argennon blockchain, and each
committee of validators generates a certificate every $m$ blocks. A validators' committee will
certify a block only if it has already been certified by the committee of delegates. Every committee of validators has
an index between $0$ and $m - 1$, and it issues a certificate for block number $n$, if $n\mod m$ equals
the committee index.

A block is considered final only if it has a valid certificate from both the delegates and the
validators. A valid certificate from validators could be either a valid certificate from the committee of
validators with the appropriate index or a certificate from all the validators.

In addition to primary committees, Argennon has several community driven committees. Certificates of these
committees are not required for block finality, but they could be used by the members of the
validators' committee to better decide about the validity of a block.

@@

The committee of delegates is a small committee of trusted delegates, elected by Argennon users either through the
Argennon Decentralized Autonomous Governance system (ADAGs\footnote{pronounced \textipa{/eI-dagz/}.}) or by emergency
voting during the recovery protocol. At the start of the Argennon mainnet, this committee will have
five members, and later its size could be changed by the ADAGs in a procedure described
in Chapter~\ref{ch:governance}.

@@

Besides the main delegates' committee, a reserve committee of delegates consists of three members is elected by users
through the ADAGs. In case the main committee fails to generate new blocks or behaves maliciously, the task of
block generation will be assigned to the reserve committee until a new main delegates' committee is elected
through the ADAGs.

@@

The Argennon protocol calculates a stake value for every account, which is an estimate of a user's stake in the
system, and is measured in ARGs. Any account whose stake value s higher than
\texttt{minValidatorsStake} threshold is considered a validator.
The \texttt{minValidatorsStake}
threshold is determined by the ADAGs, but it can never be higher than $100$ ARGs.

Every $n$ blocks randomly $m$ committees are selected from
validators, in a way that the total stake of members of each committee is approximately equal and every
account is a member of \textbf{at least} one committee.

Every member of a validators' committee has a status which can be either \texttt{online} or \texttt{offline}.
This status is stored in the ARG smart contract and is a part of the staking database. A validator can change
his status through a method invocation
from the ARG smart contract. When an account sets its status to \texttt{offline}, it receives a small reward, and
it can not change it back to \texttt{online} for \texttt{statusCoolDown} number of blocks.

@@

A block certificate issued by some members of a validators' committee is considered valid, if according to
the staking database of the previous block \textbf{certified by the same committee}, we have:
\begin{itemize}
    \item The total stake of \texttt{online} members of the committee is higher than \texttt{minOnlineStake} percent
    of the total stake of all members. This threshold can be changed by the ADAGs, but it can never be lower
    than 70 percent.
    \item All signers of the certificate have \texttt{online} status.
    \item The sum of stake values of the certificate signers is higher than 80 percent of the total stake
    of committee members that have \texttt{online} status.
\end{itemize}

If the total online stake of a committee goes lower than \texttt{minOnlineStake} threshold, a partial reshuffling
of committee members will be performed by the protocol in order to move some online members to the committee without
enough online stake. If the reshuffling can not solve the problem, nodes will initiate the recovery consensus
protocol.

As we mentioned before, every block of the Argennon blockchain needs a certificate from a specific committee of
validators. To decide about signing the block certificate, an honest member of a
validators' committee checks the conditional validity of the committee block, and if the block is valid he issues
an "accept" signature, otherwise he issues a "reject" signature.

block certificates issued by validators' committees are included in the blocks of the Argennon blockchain.
A block can contain multiple certificates, provided that those certificates belong to consecutive blocks.

@@

If a committee of validators, at any point issues a "reject" certificate or does not issue a certificate for its
block, after \texttt{certificateTimeOut} period of time,
nodes will initiate the recovery protocol.

\subsubsection{Signature Aggregation}

@@

The role of the delegates in the signature aggregation is limited. The important part of the work is done by ZK-EDB
servers. As long as there are enough honest ZK-EDB servers, the network will be able to perform signature aggregation
even if the delegates are malicious. (See Section~\ref{subsec:recovery})

@@

\subsection{The Recovery Protocol}\label{subsec:recovery}
\note{not yet written...}
%
%Validators stop issuing certificates for normal blocks and they will only issue certificates for \emph{statuts} blocks.
%A status block is a special block which can only contain status change transactions. Status blocks need to be
%certified by all the validators.
%all online validators will check the
%validity of that block. if they do not issue a certificate for that black either
\subsection{Analysis}\label{subsec:consensus-math}
\note{not yet written...}

>>>>

The credibility of a block of the Argennon blockchain is determined based on the certificates it has received
from different sets of users, knows as committees. There are two primary types of certificate committees in
Argennon: the committee of \emph{delegates} and the committee
of \emph{validators}. Argennon has one committee of delegates and $m$ committees of validators.

The committee of delegates issues a certificate for every block of the Argennon blockchain, and each
committee of validators issues a certificate every $m$ blocks. A validators' committee will
certify a block only if it has already been certified by the committee of delegates. Every committee of validators has
an index between $0$ and $m - 1$, and it issues a certificate for block number $n$, if $n$ modulo $m$ equals
the committee index.

Every block of the Argennon blockchain needs a certificate from both the committee of delegates and
the committee of validators. \textbf{A block is considered final after its next block receives both of
its certificates.}

In addition to primary committees, Argennon has several community driven committees. Certificates of these
committees are not required for block finality, but they could be used by members of the
validators' committee to better decide about the validity of a block.

@@

The committee of delegates is a small committee of trusted delegates, elected by Argennon users through the
Argennon Decentralized Autonomous Governance system (ADAGs\footnote{pronounced \textipa{/eI-dagz/}.}).
At the start of the Argennon mainnet, this committee will have
five members, and later its size could be changed by the ADAGs in a procedure described
in Section~\ref{sec:adags}.

@@

Besides the main delegates' committee, a reserve committee of delegates consists of three members is elected by users
either through the ADAGs or by \emph{emergency agreement} during the recovery protocol. In case the main committee
fails to generate new blocks or behaves maliciously, the task of
block generation will be assigned to the reserve committee until a new main delegates' committee is elected
through the ADAGs.

@@

The Argennon protocol calculates a stake value for every account, which is an estimate of a user's stake in the
system, and is measured in ARGs. Any account whose stake value is higher than
\texttt{minValidatorsStake} threshold is considered a \emph{validator}.
The \texttt{minValidatorsStake}
threshold is determined by the ADAGs, but it can never be higher than $500$ ARGs.

Every \texttt{committeeLifeTime} blocks, randomly $m$ committees are selected from
validators, in a way that the total stake of members of each committee is approximately equal, and every
account is a member of \textbf{at least} one committee.

Every validator has a status which can be either \texttt{online} or \texttt{offline}.
This status is stored in the ARG smart contract and is a part of the staking database. A validator can change
his status through a method invocation
from the ARG smart contract. When an account sets its status to \texttt{offline}, it receives a small reward, and
it can not change it back to \texttt{online} for \texttt{statusCoolDown} number of blocks.

@@

A block certificate issued by some members of a validators' committee is considered valid, if according to
the staking database of the previous block \textbf{certified by the same committee}, we have:\footnote{If we calculate
the stake values based on the previous block a malicious committee can affect the validators of the next block.}
\begin{itemize}
    \item The total stake of \texttt{online} members of the committee is higher than \texttt{minOnlineStake} percent
    of the total stake of the committee. This threshold can be changed by the ADAGs, but it can never be lower
    than $2/3$.
    \item All signers of the certificate have \texttt{online} status.
    \item The sum of stake values of certificate signers is higher than 80 percent of the total stake
    of the committee members that have \texttt{online} status.
\end{itemize}

The delegates can generate blocks very fast. Therefore, the Argennon blockchain always has an
unvalidated part which contains blocks that have a certificate from the committee of delegates but have not received
a certificate from the validators yet.

As we mentioned before, the block with height $n$ needs a certificate from the committee of
validators with index $n$ modulo $m$. To decide about signing a block certificate, a validator
checks the conditional validity\footnote{See Section~\ref{subsec:block-validation}} of the block, and
if the block is valid he issues
an "accept" signature. If the block is invalid, he initiates the recovery protocol. The validator will broadcast the
certificate \textbf{only after} he sees the certificate of the validators of the previous block.

Some validators may also require seeing a certificate from
some community driven committee for the previous block, before they broadcast an "accept" certificate for a block.

So in Argennon, the block validation by committees is performed in parallel, and validators
do not wait for seeing the validators' certificate of the previous block to start transaction validation. On the
other hand, the block certificates are published and broadcast sequentially. This ensures that an invalid
fork made by malicious delegates will not receive any certificates from validators.

@@

block certificates issued by committees of validators are included in the blocks of the Argennon blockchain.
A block can contain multiple certificates, provided that
those certificates belong to consecutive blocks.

\subsection{Status Blocks}\label{subsec:status-blocks}

If according to the staking database of block $n$, the total online stake of the committee with index $n$ modulo $m$ is
lower than \texttt{minOnlineStake} threshold, the block $n + m$ can never be certified by validators.

To prevent blockchain from halting in such situations, the protocol performs a predefined partial
reshuffling of committee members.
In this reshuffling which is based on the block random seed, some online members from other committees will
be moved to the committee without enough online stake to make it active again.

If the reshuffling can not solve the problem due to low total online stake, the protocol requires the next block
of the blockchain to be a special \emph{status} block. A status block is a special block which can only contain
status change transactions. The status block need to be certified by the delegates and by 2/3 of the total stake of
the validators. The \texttt{online}/\texttt{offline} status of validators will not be considered in the validity of
the status block certificate.\footnote{Theoretically at the status block, the total online stake of the system could
be very low, even zero. Therefore, the status block should not be certified only by online stake.} After applying
the transactions of the status block, the total online stake of validators
must go higher than \texttt{minOnlineStake} threshold.

\subsection{Signature Aggregation}\label{subsec:sig-agg}

@@

The role of the delegates in the signature aggregation is limited. The important part of the work is done by ZK-EDB
servers. As long as there are enough honest ZK-EDB servers, the network will be able to perform signature aggregation
even if the delegates are malicious.

\subsection{The Recovery Protocol}\label{subsec:recovery}

The recovery protocol is a resilient protocol intended for recovering the Argennon blockchain from critical situations.
In the terminology of the CAP theorem, the recovery protocol is designed to choose consistency over availability
and is not a protocol supposed to be executed occasionally. Ideally this protocol should never be used
during the lifetime of the Argennon blockchain.

@@

\subsection{Analysis}\label{subsec:consensus-math}
\note{not yet written...}

### 2021-06-28 tex/sections/draft.tex
Add changes needed by the recovery protocol
Signed-off-by: aybehrouz <behrouz_ayati@yahoo.com>

>>>>

If a block receives a "reject" certificate from its committee or does not receive a certificate
after \texttt{certificateTimeOut} period of time, it will be conditionally validated by \textbf{all} validators. If
it does not receive an "accept" certificate, the validators initiate the recovery protocol.

The recovery protocol can recover the functionality
of the Argennon blockchain as long as more than 2/3 of the total stake of the system is controlled by honest users
and any network partition resolves after a finite amount of time. The recovery protocol uses two main emergency
procedures to recover the : emergency forking and emergency agreement protocol.

\subsubsection{Emergency Forking}

The reserve committee of delegates can fork the Argennon blockchain, if it gets a valid fork request from
the validators. A valid fork request at
block $b$ is a signature issued by at least $0.8$ of the total \texttt{online} stake
of the validators. The stake values are calculated based on the staking database at block $b - 1$.

For forking at block $b$, the reserve committee of delegates
makes a special fork block which only contains a valid fork request at block $b$, and its parent block is block $b$.
The fork block needs a valid certificate from all $m$ validator committees in order to be valid. A valid fork block
will force the fork, in the sense that any validator receiving it considers the forked chain as the new valid chain.

A validator will issue
an "accept" certificate for a fork block at position $b + 1$ if
\begin{itemize}
    \item the fork block is signed by the reserve committee of delegates.
    \item the fork block contains a valid fork request at block $b$ and block $b$ is its parent.
    \item the validator has not issued a certificate for any block $\geq b + 2$.
    \item the validator has not seen a certificate of validators for any block $\geq b + 2$.
\end{itemize}

Emergency forking may discard at max one block with a valid validator's certificate
from the blockchain, in the worst case scenario.

\subsubsection{Emergency Agreement Protocol}

Emergency agreement protocol is a protocol for deciding between a set of proposals when no committee of
delegates is available. For initiating the protocol, a validator signs a message containing
the hash of a block which will the stake values during voting rounds will be calculated based on its
staking database. A validator starts the protocol if he
receives a message containing a valid block signed by more than 2/3 of the total stake of the validators.
if the block is too old the validator will not enter the agreement protocol.

Emergency agreement is done by voting in rounds. Each round lasts for approximately $\lambda$ units of time.
Users cast two type of votes: \emph{$i$-votes}, which are votes that are valid only in round $i$,
and \emph{final-votes}, which are votes that are valid in any round. A user executes the
following procedure in round $r$ of the voting phase:
\begin{itemize}
    \item if the user has not yet final-voted any value or if $r\mod k = 1$ he $r$-votes a desired proposal
    \item if he sees more than 2/3 round $r$-votes for a proposal $p$ he final-votes $p$
    \item if he sees more than 2/3 final-votes for a proposal he goes to the confirmation phase for $p$
    \item if $clock > r \cdot \lambda$ user goes to the round $r + 1$
\end{itemize}

the confirmation phase
\begin{itemize}
    \item user confirm-votes $p$
    \item if he sees more than 2/3 confirm votes for $p$ he selects $p$ and ends the agreement protocol
    \item if $clock > k \cdot \lambda$ user goes to the round $k + 1$ of voting
\end{itemize}

\subsubsection{Initiating the Recovery Protocol}

A validator will try to initiate the recovery protocol in two situations: when he does not receive any blocks for
\texttt{blockTimeOut} amount of time, or when he sees an evidence which proves the delegates are malicious.

When a validator is not receiving any blocks, first he switches his network to censorship resilient mode. If still he
does not receive any new blocks, he will start validating all the blocks in his blockchain that does not have
a validators' certificate. After he successfully validates the last block, if still he does not receive any blocks,
he will sign and broadcast an \textbf{emergency fork request} message at the last block he has
verified.

If the reserve committee of delegates is already active or If a validator sees a fork request signed
by 2/3 of the total stake of the validators, but does not receive the fork
block after a certain amount of time he will sign and broadcast a request for
emergency agreement on a new reserve committee. the validator puts the last valid block of his
blockchain in the request.

If a validator sees an evidence that proves the delegates are behaving maliciously, he signs and broadcasts a
fork request at approprate block. if the reserve committee is generating blocks he signs a agreement for choosing
a new reserve committee of delegates.

A validator finds a block that conditionally is not valid. in this case validators start validating previous blocks of the
invalid block that does not have a validators' certificate. then the validator sign a fork request at the last valid
block before the invalid block.
the block depends on the specific evidence
recieving a malformed block signed by the delegates for example a block which does not contain all the required feilds

two different blocks signed by the delegates at the same height (same block number)
a block signed by the delegates which is not conditionlly valid and its previous block has a valid certificate from the validators commitee

time out in receiving validators certificate

time out in receiving a new block

finds an invalid block in his blockchain

Online stake goes too low

The delegates stop generating blocks

\paragraph{The delegates fork the blockchain}
%
%Validators stop issuing certificates for normal blocks and they will only issue certificates for \emph{statuts} blocks.
%A status block is a special block which can only contain status change transactions. Status blocks need to be
%certified by all the validators.
%all online validators will check the
%validity of that block. if they do not issue a certificate for that black either
